---
title: "Data visualisation"
author: "Leah Kemp"
date: "23/08/2022"
output: html_document
---

```{r setup, include=FALSE}
# setup default chunk settings
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.align = "center")
```

```{r, eval = FALSE}
# install packages
install.packages("remotes")
remotes::install_github("robinweide/GENOVA")
install.packages("pkgfilecache")
install.packages("strawr")
install.packages("plotly")
```

```{r}
# working from the docs
# https://github.com/robinweide/GENOVA/blob/master/vignettes/GENOVA_vignette.pdf

# set project directory
project_dir <- "/NGS/scratch/KSCBIOM/HumanGenomics/guinea_pore_c/"

# load packages
library(GENOVA)
library(pkgfilecache)
library(strawr)
library(plotly)
```

```{r, eval = FALSE}
# load .hic file
hic <- GENOVA::load_contacts("/NGS/scratch/KSCBIOM/HumanGenomics/guinea_pore_c/results/04_pore_c_snakemake_pipeline_run/Pore-C-Snakemake/results/juicebox/NlaIII_run01_Cavpor3.0_unphased.hic",
                             sample_name = "guinea",
                             resolution = 10e3,
                             balancing = "KR",
                             colour = "red")

# load in centromere information
#centromeres <- utils::read.delim("./results/04_pore_c_snakemake_pipeline_run/Pore-C-Snakemake/results/juicebox/.bed",
#                                 sep = "\t",
#                                 header = FALSE,
#                                 stringsAsFactors = FALSE)
```

```{r}
# get test data
test_data <- GENOVA::get_test_data("40k")
```

## Visualisation {.tabset .tabset-fade}

### Cis-quantification

"The expected amount of intra-chromosomal contacts is in the range of 90-93%. Assuming extra inter-chromosomal contacts are due to debris/noise, the user might aspire to get the cis-percentages as close to 90% as possible." (https://github.com/robinweide/GENOVA/blob/master/vignettes/GENOVA_vignette.pdf)

```{r, out.width = "100%"}
# compute percentage of per-sample cis-contacts
cis_chrom <- GENOVA::cis_trans(test_data)

# generate barplot of percentages cis per sample
plotly::plot_ly(x = cis_chrom$sample,
                y = cis_chrom$cis,
                type = "bar",
                textposition = "none",
                name = "Raw sequencing \n read count",
                hoverinfo = "text",
                text = ~base::paste("</br> Sample:", cis_chrom$sample,
                                    "</br> cis percentage:", base::format(cis_chrom$cis, big.mark = ",", scientific = FALSE, digits = 4), "%")) %>%
  plotly::layout(xaxis = list(title = "Sample"),
                 yaxis = list(title = "cis percentage"))
```

### RCP

Relative Contact Probability (RCP), contact probability as a function of genomic distance

```{r}
# calculate the Relative Contact Probability (RCP)
# contact probability as a function of genomic distance
rcp <- GENOVA::RCP(explist = base::list(test_data),
                   chromsToUse = "chr21")

# plot
GENOVA::visualise(rcp)
```

### Compartment scores

```{r}
# read in information on the active parts of the genome
#peaks <- utils::read.delim(".bed",
#                           header = FALSE)

# calculate compartment scores using information on active parts of the genome
cs <- GENOVA::compartment_score(test_data)

# plot compartment scores on chromosome 21
GENOVA::visualise(cs, chr = "chr21")
```

### Saddle plot

```{r}
# perform quantile vs. quantile analysis of compartment score
saddle <- GENOVA::saddle(test_data,
                         CS_discovery = cs,
                         bins = 50)

# plot saddle-plot
GENOVA::visualise(saddle)
```

### Compartment strength

```{r}
# calculate compartment strength - score for how much A-A and B-B interactions occur vs. A-B interactions
css <- GENOVA::quantify(saddle)

compared <- tidyr::spread(base::unique(css[,-c(3,4)]), key = "exp", value = "strength")

with(compared, plot(Hap1_WT_40k,
                    xlim = c(0,7),
                    ylim = c(0,7),
                    pch = 20))
abline(a = 0,
       b = 1,
       lty = 1)
```

### Interaction plots

Cis compartment plot - observed over expected

```{r}
GENOVA::compartment_matrixplot(exp1 = test_data,
                               CS_discovery = cs,
                               chrom = "chr21",
                               arm = "q",
                               metric = "obsexp")
```

### Domainogram

```{r}
# calculate insulation domainogram
id <- GENOVA::insulation_domainogram(test_data,
                                     chrom = "chr21",
                                     start = 25e6,
                                     end = 29e6,
                                     window_range = c(1, 101),
                                     step = 2)

# plot
GENOVA::visualise(id)
```

### Computing the insulation score

```{r}
# calculate the genome wide insulation score
insulation <- GENOVA::insulation_score(test_data,
                                       window = 25)

# plot
GENOVA::visualise(insulation,
                  chr = "chr21",
                  start = 25e6,
                  end = 29e6)
```

